---
title: "gplus"
output: html_document
---

```{r}
library(tidyverse)
library(devtools)
library(ggplot2)
library(remotes)
library(SBpitch)
library(StatsBombR)
library(grid)
library(data.table)
library(tidyr)
library(worldfootballR)
```

```{r}
Comp <- FreeCompetitions()
Matches <- FreeMatches(Comp)
```

```{r}
"DO NOT USE"
#matches_1516 <- Matches %>%
  #filter(competition.competition_name == "1. Bundesliga" | competition.competition_name ==  "La Liga" | competition.competition_name == "Ligue 1" | competition.competition_name == "Serie A" | competition.competition_name == "Premier League") %>%
  #filter(season.season_name == "2015/2016")
```

```{r}
germany_matches_1516 <- Matches %>%
  filter(competition.competition_name == "1. Bundesliga") %>%
  filter(season.season_name == "2015/2016")

germany_events_1516 <- free_allevents(MatchesDF = germany_matches_1516, Parallel = T)
germany_events_1516 <- allclean(germany_events_1516)
germany_events_1516 <- get.opposingteam(germany_events_1516)


italy_matches_1516 <- Matches %>%
  filter(competition.competition_name == "Serie A") %>%
  filter(season.season_name == "2015/2016")

italy_events_1516 <- free_allevents(MatchesDF = italy_matches_1516, Parallel = T)
italy_events_1516 <- allclean(italy_events_1516)
italy_events_1516 <- get.opposingteam(italy_events_1516)


england_matches_1516 <- Matches %>%
  filter(competition.competition_name == "Premier League") %>%
  filter(season.season_name == "2015/2016")

england_events_1516 <- free_allevents(MatchesDF = england_matches_1516, Parallel = T)
england_events_1516 <- allclean(england_events_1516)
england_events_1516 <- get.opposingteam(england_events_1516)

spain_matches_1516 <- Matches %>%
  filter(competition.competition_name == "La Liga") %>%
  filter(season.season_name == "2015/2016")

spain_events_1516 <- free_allevents(MatchesDF = spain_matches_1516, Parallel = T)
spain_events_1516 <- allclean(spain_events_1516)
spain_events_1516 <- get.opposingteam(spain_events_1516)

france_matches_1516 <- Matches %>%
  filter(competition.competition_name == "Ligue 1") %>%
  filter(season.season_name == "2015/2016")

france_events_1516 <- free_allevents(MatchesDF = france_matches_1516, Parallel = T)
france_events_1516 <- allclean(france_events_1516)
france_events_1516 <- get.opposingteam(france_events_1516)
```

```{r}
# write.csv(all_events_1516,"/Users/KIDS/Downloads/1516_data_events.csv")
# write.csv(germany_events_1516,"/Users/KIDS/Downloads/1516_germany_events.csv")
# write.csv(italy_events_1516,"/Users/KIDS/Downloads/1516_italy_events.csv")
# write.csv(england_events_1516,"/Users/KIDS/Downloads/1516_england_events.csv")
# write.csv(spain_events_1516,"/Users/KIDS/Downloads/1516_spain_events.csv")
# write.csv(france_events_1516,"/Users/KIDS/Downloads/1516_france_events.csv")
```

```{r}
all_events_1516 <- bind_rows(germany_events_1516,spain_events_1516,england_events_1516,france_events_1516,italy_events_1516)

all_events_1516 <- all_events_1516 %>%
  mutate(minsec = 60*minute + second)

germany_events_1516 <- germany_events_1516 %>%
  mutate(minsec = 60*minute + second)

england_events_1516 <- england_events_1516 %>%
  mutate(minsec = 60*minute + second)

italy_events_1516 <- italy_events_1516 %>%
  mutate(minsec = 60*minute + second)

france_events_1516 <- france_events_1516 %>%
  mutate(minsec = 60*minute + second)

spain_events_1516 <- spain_events_1516 %>%
  mutate(minsec = 60*minute + second)

saveRDS(all_events_1516, "/Users/KIDS/Downloads/all_events_1516.rds")
saveRDS(germany_events_1516, "/Users/KIDS/Downloads/germany_events_1516.rds")
saveRDS(england_events_1516, "/Users/KIDS/Downloads/england_events_1516.rds")
saveRDS(italy_events_1516, "/Users/KIDS/Downloads/italy_events_1516.rds")
saveRDS(france_events_1516, "/Users/KIDS/Downloads/france_events_1516.rds")
saveRDS(spain_events_1516, "/Users/KIDS/Downloads/spain_events_1516.rds")
```

```{r}
"USE TO RESTORE"

all_events_1516 <- readRDS("/Users/KIDS/Downloads/all_events_1516.rds")

# germany_events_1516 <- readRDS("/Users/KIDS/Downloads/germany_events_1516.rds")
# italy_events_1516 <- readRDS("/Users/KIDS/Downloads/italy_events_1516.rds")
# england_events_1516 <- readRDS("/Users/KIDS/Downloads/england_events_1516.rds")
# france_events_1516 <- readRDS("/Users/KIDS/Downloads/france_events_1516.rds")
# spain_events_1516 <- readRDS("/Users/KIDS/Downloads/spain_events_1516.rds")
```

```{r}
all_events_1516_selectcols <- all_events_1516 %>%
  select(id,index,period,under_pressure,off_camera,type.name,possession,possession_team.name,team.name,pass.length,pass.angle,pass.cross,pass.cut_back,pass.recipient.name,pass.height.name,pass.body_part.name,pass.type.name,pass.outcome.name,pass.technique.name,ball_receipt.outcome.name,interception.outcome.name,shot.statsbomb_xg,duel.type.name,duel.outcome.name,dribble.outcome.name,clearance.body_part.name,match_id,foul_committed.type.name,location.x,location.y,minsec,carry.end_location.x,carry.end_location.y,pass.end_location.x,pass.end_location.y,OpposingTeam, StartOfPossession, TimeInPoss, TimeToPossEnd, position.name, match_id,pass.type.name,player.name)
```

```{r}
head(all_events_1516,25)
```

NEXT STEPS:

 - create new all_events df with only meaningful predictors
    - create necessary cols like prior pass longball/cross/etc binaries
    - delete things as needed (e.g. maybe merge receipt into pass)
 - xgboost it
 
 Maybe keep ball receipts in  but ensure pass g+ isn't double counted in total calculation
  - Or find a good way to split g+ for passes between passer and receiver (xpass!!!)
 
Make a list of things that shouldn't be counted in event count and implement function to not count them
  - Think about whether ball receipts should be events (ask asa?)]


```{r}
table(all_events_1516$pass.body_part.name)
print("x")
table(all_events_1516$shot.body_part.name)
print("x")
table(all_events_1516$clearance.body_part.name)
print("x")
table(all_events_1516$goalkeeper.body_part.name)
print("x")

table(all_events_1516$type.name) #figure out which are "real" events - filter and merge accordingly to create a cohesive gplus df
```

```{r}
"THROW INS"
total_xg = 0
total_shots = 0
total_throws = 0

for (x in 1:nrow(all_events_1516_selectcols)) {
  
   if(x %% 100000 == 0) {
    print(100*x/nrow(all_events_1516_selectcols))
   }
  
  pass_type_name <- all_events_1516_selectcols$pass.type.name[x]
  location_x <- all_events_1516_selectcols$location.x[x]
  off_camera <- all_events_1516_selectcols$off_camera[x]
  
  if (!is.na(pass_type_name) && !is.na(location_x) && is.na(off_camera)) {
    if (pass_type_name == "Throw-in" && location_x > 80) {
      total_throws = total_throws + 1
      indiv_xg = 0
      event_row = all_events_1516_selectcols[x, ]
      
      for (n in 1:30) {
        next_idx <- x + n
        if (next_idx <= nrow(all_events_1516_selectcols)) {
          if (all_events_1516_selectcols$minsec[next_idx] < all_events_1516_selectcols$minsec[x] + 10 &&
              all_events_1516_selectcols$match_id[next_idx] == all_events_1516_selectcols$match_id[x]) {
            event_row <- rbind(event_row, all_events_1516_selectcols[next_idx, ])
          }
        }
      }
      
      for (a in 1:nrow(event_row)) {
        shot_xg <- event_row$shot.statsbomb_xg[a]
        if (!is.na(shot_xg)) {
          total_shots = total_shots + 1
          total_xg = total_xg + shot_xg
        }
      }
    }
  }
}

cat("Throw-ins in final third: ", paste(total_throws), "\n")
cat("Shots from these throw-ins: ", paste(total_shots), "\n")
cat("xG per shot: ", paste(total_xg/total_shots), "\n")
```

```{r}
"PASSES (COMPLETE (add this!), shots same team as passes - no conceded xg yet)"

total_xg = 0
total_shots = 0
total_passes = 0

#the box to be drawn around (x_start, y_start) with sides 2*radius_start - might be too rigid 
x_start = 95
y_start = 40
radius_start = 5
  
x_end = 110
y_end = 40
radius_end = 5
  
for (x in 1:nrow(all_events_1516_selectcols)) {
  
   if(x %% 500000 == 0) {
    print(100*x/nrow(all_events_1516_selectcols))
   }
  
  pass_type_name <- all_events_1516_selectcols$pass.type.name[x]
  location_x <- all_events_1516_selectcols$location.x[x]
  location_y <- all_events_1516_selectcols$location.y[x]
  end_location_x <- all_events_1516_selectcols$pass.end_location.x[x]
  end_location_y <- all_events_1516_selectcols$pass.end_location.y[x]
  off_camera <- all_events_1516_selectcols$off_camera[x]
  
  
  if (!is.na(pass_type_name) && !is.na(location_x) && is.na(off_camera)) {
    if (location_x > x_start - radius_start && location_x < x_start + radius_start & 
        location_y > y_start - radius_start && location_y < y_start + radius_start &
        end_location_x > x_end - radius_end && end_location_x < x_end + radius_end &
        end_location_y > y_end - radius_end && end_location_y < y_end + radius_end) {
      
      total_passes = total_passes + 1
      indiv_xg = 0
      event_row = all_events_1516_selectcols[x, ]
      
      for (n in 1:30) {
        next_idx <- x + n
        if (next_idx <= nrow(all_events_1516_selectcols)) {
          if (all_events_1516_selectcols$minsec[next_idx] < all_events_1516_selectcols$minsec[x] + 10 &&
              all_events_1516_selectcols$match_id[next_idx] == all_events_1516_selectcols$match_id[x] &&
              all_events_1516_selectcols$team.name[next_idx] == all_events_1516_selectcols$team.name[x]) {
            event_row <- rbind(event_row, all_events_1516_selectcols[next_idx, ])
          }
        }
      }
      
      for (a in 1:nrow(event_row)) {
        shot_xg <- event_row$shot.statsbomb_xg[a]
        if (!is.na(shot_xg)) {
          total_shots = total_shots + 1
          total_xg = total_xg + shot_xg
        }
      }
    }
  }
}

cat("Passes in range: ", paste(total_passes), "\n")
cat("Shots within 10 seconds from these passes: ", paste(total_shots), "\n")
cat("xG per shot: ", paste(total_xg/total_shots), "\n")
cat("xG per pass: ", paste(total_xg/total_passes))
```

```{r}
create_Pitch() +   
  geom_rect(aes(xmin = x_start-radius_start, xmax = x_start+radius_start, ymin = y_start-radius_start, ymax = y_start+radius_start),
            fill = "red", alpha = 0.3) +
  geom_rect(aes(xmin = x_end-radius_end, xmax = x_end+radius_end, ymin = y_end-radius_end, ymax = y_end+radius_end),
            fill = "red", alpha = 0.3) +
  geom_point(aes(x=x_start,y=y_start)) + 
  geom_segment(aes(x=x_start,y=y_start,xend=x_end,yend=y_end),alpha=0.75,arrow=arrow(length=unit(0.2,"cm"))) + 
  theme(plot.title = element_text(size = 24, color = "black",family="sans",face="bold", hjust = 0.5,vjust=-2),
        plot.subtitle = element_text(size = 12, color = "black", family = "sans", hjust = 0.5,vjust = -4),
        plot.caption = element_text(size = 12, color = "black", family = "sans", hjust = 0.5,vjust = 4)) + 
  scale_y_reverse() +
  labs(title = paste("Shots per pass:", round(total_shots/total_passes,3)), 
       subtitle = paste("xG per pass:", round(total_xg/total_passes,3), "|", "Passes:", total_passes), 
       caption = "Caption")

```


```{r}
"POSSESSION VALUE AT GIVEN (X,Y) - BALL RECEIPTS"

total_xg_for = 0
total_shots_for = 0

total_xg_against = 0
total_shots_against = 0

total_receipts = 0

x_receipt = 20
y_receipt = 40
radius_receipt = 5

for (x in 1:nrow(all_events_1516_selectcols)) {
  
   if(x %% 1500000 == 0) {
    print(100*x/nrow(all_events_1516_selectcols))
   }
  
  location_x <- all_events_1516_selectcols$location.x[x]
  location_y <- all_events_1516_selectcols$location.y[x]
  off_camera <- all_events_1516_selectcols$off_camera[x]
  type_name <- all_events_1516_selectcols$type.name[x]
  poss_number <- all_events_1516_selectcols$possession[x] #easier references
  team_for <- all_events_1516_selectcols$team.name[x]
  team_against <- all_events_1516_selectcols$OpposingTeam[x]
  
  if (!is.na(location_x) && !is.na(location_y) && is.na(off_camera)) {
    if (location_x > x_receipt - radius_receipt && location_x < x_receipt + radius_receipt & #if it's a ball receipt in the given area
        location_y > y_receipt - radius_receipt && location_y < y_receipt + radius_receipt &
        type_name == "Ball Receipt*") {
      
      total_receipts = total_receipts + 1
      event_row = all_events_1516_selectcols[x, ]
      
      for (n in 1:30) { 
        next_idx <- x + n
        if (next_idx <= nrow(all_events_1516_selectcols)) {
          if (all_events_1516_selectcols$minsec[next_idx] < all_events_1516_selectcols$minsec[x] + 60 && #less than a minute later 
              all_events_1516_selectcols$match_id[next_idx] == all_events_1516_selectcols$match_id[x] && #same game
              #all_events_1516_selectcols$team.name[next_idx] == all_events_1516_selectcols$team.name[x] && #same team
              all_events_1516_selectcols$possession[next_idx] <= all_events_1516_selectcols$possession[x] + 1) { #same or next possession
            event_row <- rbind(event_row, all_events_1516_selectcols[next_idx, ])
          }
        }
      }
      
      for (a in 1:nrow(event_row)) {
        shot_xg <- event_row$shot.statsbomb_xg[a]
        if (!is.na(shot_xg)) {
          if (event_row$team.name[a] == team_for) {
            total_shots_for = total_shots_for + 1
            total_xg_for = total_xg_for + shot_xg
          }
          else if (event_row$team.name[a] == team_against) {
            total_shots_against = total_shots_against + 1
            total_xg_against = total_xg_against + shot_xg
          }
        }
      }
    }
  }
}

cat("Receipts: ", paste(total_receipts), "\n")
cat("Shots for: ", paste(total_shots_for), "\n")
cat("xG for: ", paste(total_xg_for), "\n")
cat("Shots against: ", paste(total_shots_against), "\n")
cat("xG against: ", paste(total_xg_against), "\n")
```



```{r}
create_Pitch() +   
  geom_rect(aes(xmin = x_receipt-radius_receipt, xmax = x_receipt+radius_receipt, ymin = y_receipt-radius_receipt, ymax = y_receipt+radius_receipt),
            fill = "green", alpha = 0.3) +
  theme(plot.title = element_text(size = 24, color = "black",family="sans",face="bold", hjust = 0.5,vjust=-2),
        plot.subtitle = element_text(size = 12, color = "black", family = "sans", hjust = 0.5,vjust = -4),
        plot.caption = element_text(size = 12, color = "black", family = "sans", hjust = 0.5,vjust = 4)) + 
  scale_y_reverse() +
  labs(title = paste("Possessions:", total_receipts), 
       subtitle = paste("xG for per poss:", round(total_xg_for/total_receipts,5), "|", "xG against per poss:", round(total_xg_against/total_receipts,5)), 
       caption = "Caption")
```

```{r}
"PITCH_ZONES CREATION - DO NOT USE AGAIN!!!"

y_min_prelim <- seq(0, 72, by = 8)
y_min_list <- rep(y_min_prelim, 15)

y_max_prelim <- seq(8, 80, by = 8)
y_max_list <- rep(y_max_prelim, 15)

x_min_prelim <- seq(0, 112, by = 8)
x_min_list <- rep(x_min_prelim, each = 10)

x_max_prelim <- seq(8, 120, by = 8)
x_max_list <- rep(x_max_prelim, each = 10)

labels_list <- seq(1,150, by = 1)

#pitch_zones <- data.frame(
#  zone_num = labels_list,
#  x_min = x_min_list, 
#  x_max = x_max_list,
#  y_min = y_min_list, 
#  y_max = y_max_list
#)

#pitch_zones$receipts <- NA
#pitch_zones$xg_for <- NA
#pitch_zones$xg_against <- NA
#pitch_zones$shots_for <- NA
#pitch_zones$shots_against <-NA
```

```{r}
"TESTING VISUAL"
zones_ref <- create_Pitch() + 
  geom_rect(data = pitch_zones, aes(xmin = x_min, ymin = y_min, xmax = x_max, ymax = y_max, fill = xg_diff_per_rec), color = alpha("black", 0.3)) +
  scale_fill_gradient2(low = alpha("red", 1), high = alpha("darkgreen", 1), mid = alpha("white", 0.3), midpoint = 0.015) + 
  geom_text(data = pitch_zones, aes(x = (x_min + x_max) / 2, y = (y_min + y_max) / 2, label = zone_num), color = alpha("black",065), size = 3) +
  theme_void() + 
  scale_y_reverse()

zones_ref
```









































```{r}
"RUNNING EACH SQUARE - DONE!"
for (rownum in 1:nrow(pitch_zones)) {
  
  print(rownum)
  
  total_xg_for = 0
  total_shots_for = 0

  total_xg_against = 0
  total_shots_against = 0

  total_receipts = 0
  
  zone_xmin <- pitch_zones$x_min[rownum]
  zone_xmax <- pitch_zones$x_max[rownum]
  zone_ymin <- pitch_zones$y_min[rownum]
  zone_ymax <- pitch_zones$y_max[rownum]
  
  for (x in 1:nrow(all_events_1516_selectcols)) {
    
  
    location_x <- all_events_1516_selectcols$location.x[x]
    location_y <- all_events_1516_selectcols$location.y[x]
    off_camera <- all_events_1516_selectcols$off_camera[x]
    type_name <- all_events_1516_selectcols$type.name[x]
    poss_number <- all_events_1516_selectcols$possession[x] #easier references
    team_for <- all_events_1516_selectcols$team.name[x]
    team_against <- all_events_1516_selectcols$OpposingTeam[x]

    if (!is.na(location_x) && !is.na(location_y) && is.na(off_camera)) {
      if (location_x > zone_xmin && location_x < zone_xmax & #if it's a ball receipt in the given area
          location_y > zone_ymin && location_y < zone_ymax &
          type_name == "Ball Receipt*") {
      
        total_receipts = total_receipts + 1
        event_row = all_events_1516_selectcols[x, ]
      
        for (n in 1:30) { 
          next_idx <- x + n
          if (next_idx <= nrow(all_events_1516_selectcols)) {
            if (all_events_1516_selectcols$minsec[next_idx] < all_events_1516_selectcols$minsec[x] + 60 && #less than a minute later 
                all_events_1516_selectcols$match_id[next_idx] == all_events_1516_selectcols$match_id[x] && #same game
                #all_events_1516_selectcols$team.name[next_idx] == all_events_1516_selectcols$team.name[x] && #same team
                all_events_1516_selectcols$possession[next_idx] <= all_events_1516_selectcols$possession[x] + 1) { #same or next possession
              event_row <- rbind(event_row, all_events_1516_selectcols[next_idx, ])
            }
          }
        }
      
        for (a in 1:nrow(event_row)) {
          shot_xg <- event_row$shot.statsbomb_xg[a]
          if (!is.na(shot_xg)) {
            if (event_row$team.name[a] == team_for) {
              total_shots_for = total_shots_for + 1
              total_xg_for = total_xg_for + shot_xg
            }
            else if (event_row$team.name[a] == team_against) {
              total_shots_against = total_shots_against + 1
              total_xg_against = total_xg_against + shot_xg
            }
          }
        }
      }
    }
  }

  cat("Receipts: ", paste(total_receipts), "\n")
  cat("Shots for: ", paste(total_shots_for), "\n")
  cat("xG for: ", paste(total_xg_for), "\n")
  cat("Shots against: ", paste(total_shots_against), "\n")
  cat("xG against: ", paste(total_xg_against), "\n")
  
  pitch_zones$receipts[rownum] = total_receipts
  pitch_zones$xg_for[rownum] = total_xg_for
  pitch_zones$xg_against[rownum]  = total_xg_against
  pitch_zones$shots_for[rownum]  = total_shots_for
  pitch_zones$shots_against[rownum] <-total_shots_against
}

```

```{r}
pitch_zones <- pitch_zones %>%
  mutate(xg_diff = xg_for - xg_against, xg_diff_per_rec = (xg_for-xg_against)/receipts)
```


MAKE SURE TO WRITE CSV FOR PITCH_ZONES AS BACKUP!!!!!!

```{r}
write.csv(pitch_zones,"/Users/KIDS/Downloads/pitchzonesgplusFINAL.csv")
```


```{r}
head(all_events_1516)
```

```{r}
throwcols <- all_events_1516 %>%
  select(id,index,period,under_pressure,off_camera,type.name,possession,possession_team.name,team.name,pass.length,pass.height.name,pass.type.name,pass.outcome.name,shot.statsbomb_xg,match_id,location.x,location.y,minsec,pass.end_location.x,pass.end_location.y, position.name, match_id,pass.type.name,player.name)
```



























```{r}
throws_18 = 0

shots_for_18 = 0
xg_for_18 = 0

shots_ag_18 = 0
xg_ag_18 = 0

for(i in 1:nrow(throwcols)) {
  if(!is.na(throwcols$pass.type.name[i]) & !is.na(throwcols$location.x[i])) {
    
    if(throwcols$pass.type.name[i] == "Throw-in" & throwcols$location.x[i] < 18) { #if throw-in in own quarter
    
    throws_18 = throws_18 + 1
    
    newdf <- throwcols %>%
      filter(match_id == throwcols$match_id[i] & minsec < throwcols$minsec[i] + 120 & possession %in% c(throwcols$possession[i], throwcols$possession[i] + 1)) #new df with events in same game in next 60 secs #can delete minsec if poss works
    
    for(q in 1:nrow(newdf)) {
      if(!is.na(newdf$shot.statsbomb_xg[q])) { #if shot
        if (newdf$team.name[q] == throwcols$team.name[i]) {
          shots_for_18 = shots_for_18 + 1
          xg_for_18 = xg_for_18 + newdf$shot.statsbomb_xg[q]
        }
        else{
          shots_ag_18 = shots_ag_18 + 1
          xg_ag_18 = xg_ag_18 + newdf$shot.statsbomb_xg[q]
        }
      }
    }
  }
  if (i %% 10000 == 0) {
    print(100*i/nrow(throwcols))
  }

  }
}
```

```{r}
kickoffs = 0

shots_for_kos = 0
xg_for_kos = 0

shots_ag_kos = 0
xg_ag_kos = 0

for(i in 1:nrow(throwcols)) {
  if(!is.na(throwcols$pass.type.name[i]) & !is.na(throwcols$location.x[i])) {
    
    if(throwcols$pass.type.name[i] == "Kick Off") { #if kickoff
    
    kickoffs = kickoffs + 1
    
    newdf <- throwcols %>%
      filter(match_id == throwcols$match_id[i] & minsec < throwcols$minsec[i] + 120 & possession %in% c(throwcols$possession[i], throwcols$possession[i] + 1)) #new df with events in same game in next 60 secs #can delete minsec if poss works
    
    for(q in 1:nrow(newdf)) {
      if(!is.na(newdf$shot.statsbomb_xg[q])) { #if shot
        if (newdf$team.name[q] == throwcols$team.name[i]) {
          shots_for_kos = shots_for_kos + 1
          xg_for_kos = xg_for_kos + newdf$shot.statsbomb_xg[q]
        }
        else{
          shots_ag_kos = shots_ag + 1
          xg_ag_kos = xg_ag_kos + newdf$shot.statsbomb_xg[q]
        }
      }
    }
  }

  }
  if (i %% 100000 == 0) {
    print(100*i/nrow(throwcols))
  }
}
```


